---
title: "Final Project Code"
subtitle:  "Cheeson Lau, Edison Lu, and Noah McMahon"
graphics: yes
output: pdf_document
header-includes:
    - \usepackage{amsmath, amssymb}
    - \usepackage{framed}\definecolor{shadecolor}{rgb}{0.949,0.949,0.949}
---


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(janitor)
```

### Creating the Dataset

```{r Creating the Dataset}

cov <- tibble("Group" = c("BNT162b2", "Placebo", "Total"), 
              "COVID-19 Cases" = c(8, 162, 170), 
              "No. of Subjects" = c(17411, 17511, 34922))

```

### Checking Values from Paper

We know that P(COVID | BNT162b2) = $\pi_v$, and that P(COVID | Placebo) = $\pi_p$.

```{r pi_v and pi_p}

pi_v <- (8 / 17411)
pi_p <- (162 / 17511)

```

Looking at $T \sim Binom(n = 170, \pi)$, we know that $\pi = \frac{\pi_v}{\pi_v + \pi_p}$ from the instructions.
Additionally, the vaccine efficacy $\psi = \frac{1 - 2\pi}{1 - \pi}$.

```{r pi and psi}

pi <- (pi_v) / (pi_v + pi_p)
psi <- (1 - (2 * pi)) / (1 - pi)

```

### Maximum Likelihood Estimator

We know that $T \sim Binom(n = 170, \pi)$ and that we have observed t = 8. Additionally, the parameter of interest is the vaccine efficacy $\psi = \frac{1 - 2\pi}{1 - \pi}$.

Using the Invariance of the MLE (which we proved in Problem 6), we can say that
\begin{center}
$\hat{\psi_0}^{MLE} = \frac{1 - 2\hat{\pi_0}^{MLE}}{1 - \hat{\pi_0}^{MLE}}$
\end{center}

### Method of Moments Estimator

We know that $T \sim Binom(n = 170, \pi)$ and that we have observed t = 8. Additionally, the parameter of interest is the vaccine efficacy $\psi = \frac{1 - 2\pi}{1 - \pi}$. We know that $\pi = \frac{1 - \psi}{2 - \psi}$.

The method of moments estimator satisfies
\begin{align}
E[T] &= \bar{t} \\
170 * \pi &= \frac{8}{1} \\
170 * \frac{1 - \psi}{2 - \psi} &= 8 \\
\frac{1 - \psi}{2 - \psi} &= \frac{8}{170} \\
1 - \psi &= \frac{16}{170} - \frac{8}{170}\psi \\
\frac{154}{170} &= \frac{162}{170}\psi \\
154 &= 170\psi \\
\hat{\psi_0}^{MOM} &= \frac{154}{170}
\end{align}

We must now verify that $\hat{\psi_0}^{MOM} = \frac{n - 2T}{n + T}$.
\begin{align}
\hat{\psi_0}^{MOM} &= \frac{n - 2T}{n + T} \\
&= \frac{170 - 2T}{170 + T}
\end{align}
